---
title: "GRIP#01"
author: "Shamie Dasgupta"
date: "11 september 2022"
output: html_document
---

## task1: Predicting the percentage of a student on the number of study hours (supervised ML)

We are to build a regression model which will predict the percentage of a student on the number of study hours. 
Here Scores refers to the percentage of marks obtained by the student and Hours refers to the number of hours a student studies in a day. 

 Scores is the response variable or dependent variable and Hours is the explanatory variable or the independent variable. 

```{r}
rm(list=ls())
study=read.csv("student.csv")
attach(study)
study
```
The data is retrieved from the csv file and stored in the 'study' data frame and attached.

Now we perform some exploratory Data Analysis to know more about the data:
```{r}
#importing all the libraries required:
library(ggplot2)

```

```{r}
ggplot(study,aes(Hours))+geom_boxplot(fill='sky blue')
ggplot(study,aes(Scores))+geom_boxplot(fill=' pink')
```

Next, we plot the scatterplot to observe the relation between the variables Scores and Hours:

```{r}

ggplot(study,aes(Hours,Scores))+geom_point(col=2)+
  labs(title="Scatterplot of Scores vs Hours")
```
From the scatterplot we can see that the scores of a student is more or less positively correlated with the hours of study. Moreover the relation between them looks more or less linear. 

Based on the observations from the scatterplot we proceed to fit a linear regression model.

## The linear regression model:

Now,we divide the data into two parts randomly- referring one to train data and the other to test data. We divide the data into 3:1 ratio.

```{r}
set.seed(123)
rsam=sample(x=c('train','test'),size=nrow(study),replace=T, prob= c(0.75,0.25))
train=study[rsam=='train',]
test=study[rsam=='test',]
```

Now we fit a linear regression model to the train data:

```{r}
stu_lin=lm(Scores~Hours,train)
summary(stu_lin)
```


With the help of R, the linear regression model has come out to be :

      Scores=3.11 + 9.70Hours

the coefficient of determination for the above regression model is 0.9601.
That is, the above linear regression model explains approximately 96% of the total variation in Scores due to the number of hours of studying by a student. 
Since the coefficient of determination is very high we can say that the fit is good.

Here is the fitted model:
```{r}
ggplot(study,aes(Hours,Scores))+geom_point(col=2)+
  labs(title="Scatterplot of Scores vs Hours")+
  geom_smooth(method='lm')
```


Now we test this model on our test data:
```{r}

pred_lin <- predict(stu_lin,data.frame(Hours=test$Hours))
ggplot(NULL,aes(x=test$Hours))+
  geom_point(aes(y=test$Scores,color='red'))+
  geom_line(aes(y=pred_lin)) +geom_point(aes(y=pred_lin,color='blue'))+
  labs(x='hours in test data',y='scores')+
  scale_color_discrete(name="data points",
                         labels=c("predicted value","actual value"))
```
```{r}

error=test$Scores-pred_lin
r.sq=1- sum(error^2)/sum((test$Scores-mean(test$Scores))^2)
r.sq
```
the coefficient of determination for the above regression model when fitted on the test data is 0.9166. 
Since the coefficient of determination is still high for the test data we can say that the model is a good fit for prediction.


Now we are to find the predicted score of a student who studies 9.5 hours per day. We use the regression model to find the answer:
```{r}
hr=9.5
sc=stu_lin$coefficients[[1]] + stu_lin$coefficients[[2]]*hr;sc

```
That is, if a student studies 9.5 hours per day then he/she can be expected to score approximately 95.25% in the exam.

With that the first task of my data science and business analytics internship is completed.


